Unimodular covers of 3-dimensional parallelepipeds and Cayley sums

### 2019-07-25 unimodular-covers.tex
minor edits

Oda \cite{Oda1997} posed several questions regarding smoothness and the IDP property for lattice polytopes.
%
Following \cite{HaaseHof, Tsuchiya}, we say that a pair $(P, Q)$ of lattice polytopes has the integer decomposition property, or that \emph{the pair $(P,Q)$ is IDP}, if 
\begin{align*}
\label{eq:mixedIDP}
(P+Q) \cap \Z^d = P \cap \Z^d + Q \cap \Z^d.
\end{align*}
%
A lattice polytope $Q$ is called \emph{smooth} if it is simple and the primitive edge directions at every vertex form a linear basis for the lattice; equivalently, if the projective toric variety defined by the normal fan of $Q$ is smooth. 
The following versions of Oda's questions are now considered conjectures~\cite{HNPS2008,mfo2007}, and they are open even in dimension three:
\begin{conjecture}
\label{conj:Oda}
\begin{enumerate}
\item 
\label{itm:smoothIDP}
(Related to problems 2 and 5 in \cite{Oda1997})
Every smooth lattice polytope is IDP.
\item 
\label{itm:mixedIDP}
(Related to problems 1, 3, 4, 6 in \cite{Oda1997}) Every pair $(P,Q)$ of lattice polytopes with $Q$ smooth and the normal fan of $Q$ refining that of $P$ is IDP.
%\begin{align}
%\label{eq:mixedIDP}
%(P+Q) \cap \Z^d = P \cap \Z^d + Q \cap \Z^d.
%\end{align}
\end{enumerate}
\end{conjecture}

@@

Observe that for every $i\in \{1,\dots,k-1\}$ the intersection $P\cap(\R^2\times\{i\})$ has the same normal fan as $Q_1+Q_2$. Thus, each slice
\[
P \cap (\R^2\times[i-1,i])
\]
is a Cayley polytope. If $i\in\{2,\dots,k-1\}$, both bases have the same normal fan (and therefore each is a weak Minkowski summand of the other), while for $i\in \{1,k\}$ one base is a weak Minkowski summand of the other. We can therefore apply \Cref{thm:cayley} to each slice and combine the covers thus obtained to get a unimodular cover of $P$.
\end{proof}

>>>>

Oda (\cite{Oda1997}) posed several questions regarding smoothness and the IDP property for lattice polytopes.
%
Following \cite{HaaseHof, Tsuchiya}, we say that a pair $(P, Q)$ of lattice polytopes has the integer decomposition property, or that \emph{the pair $(P,Q)$ is IDP}, if 
\begin{align*}
\label{eq:mixedIDP}
(P+Q) \cap \Z^d = P \cap \Z^d + Q \cap \Z^d.
\end{align*}
%
A lattice polytope $Q$ is called \emph{smooth} if it is simple and the primitive edge directions at every vertex form a linear basis for the lattice; equivalently, if the projective toric variety defined by the normal fan of $Q$ is smooth. 
The following versions of Oda's questions are now considered conjectures~\cite{HNPS2008,mfo2007}, and they are open even in dimension three:
\begin{conjecture}
\label{conj:Oda}
\begin{enumerate}
\item 
\label{itm:smoothIDP}
(Related to problems 2 and 5 in \cite{Oda1997})
Every smooth lattice polytope is IDP.
\item 
\label{itm:mixedIDP}
(Related to problems 1, 3, 4, 6 in \cite{Oda1997}) Every pair $(P,Q)$ of lattice polytopes with $Q$ smooth and the normal fan of $Q$ refining that of $P$ is IDP.
%\begin{align}
%\label{eq:mixedIDP}
%(P+Q) \cap \Z^d = P \cap \Z^d + Q \cap \Z^d.
%\end{align}
\end{enumerate}
\end{conjecture}

@@

Observe that for every $i\in \{1,\dots,k-1\}$ the intersection $P\cap(\R^2\times\{i\})$ has the same normal fan as $Q_1+Q_2$. Thus, each slice
\[
P \cap (\R^2\times[i-1,i])
\]
is a Cayley polytope. For $i\in\{2,\dots,k-1\}$,  both bases have the same normal fan (and therefore each is a weak Minkowski summand of the other); for $i\in \{1,k\}$ one base is a weak Minkowski summand of the other. We can therefore apply \Cref{thm:cayley} to each slice and combine the covers thus obtained to get a unimodular cover of $P$.
\end{proof}
